Experiment 1

### 2021-01-14 C.analysis/markdowns/Experiment_1.Rmd

---
title: "Experiment 1"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook: 
    toc: true
    toc_float: true
---

@@

In experiment 1 we apply the different FoRTE girdling treatments to the ED forest stand. This stand starts from near bare ground starting in 1900. After a series of tests, we had to exclude pines from the ED configuration.

See [GitHub issue](https://github.com/FoRTExperiment/FoRTE-mgmt/issues/77#issue-687269729) for some more details.

Observations

* NPP winter pattern is interesting
* By 2030 NPP hasn't recovered 
* There is a small pattern in the Rh

Questions

* Based on discussion with Lisa do we want to take a look at aboveground bark biomass? Would that be something easy to compare with Max's dendrology band measurements? 
* Would it make sense to report values as difference harvest 45% - harvest 0%? Or the difference from a refernce value for a single year? 
* Should we use ED output to calculate resillence? 
* Should we run ED with climatological inputs?

@@

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# The defaults for the chunks
knitr::opts_chunk$set(echo = TRUE, fig.width = 5, fig.height = 3)

library(ggplot2)
library(magrittr)
library(data.table)
library(cowplot)
library(lubridate)
library(purrr)

BASE_DIR <- here::here()
THEME    <- theme_bw(base_size = 15)
```

@@

```{r}
# Note that this is the object created with the drake frame work, there should be some better way to trigger this. 
exp_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', "exp-1.rds"))
exp_object <- map(exp_object, function(x){ x[year < 2030, ] }) # Note the year 2030 is not a complete run 
```

@@

@@

```{r, fig.width = 3, fig.height = 3}
ggplot() + 
  geom_point(data = exp_object$NPP[scn == 'harvest_0', ], aes(year, value)) +
  THEME +
  labs(title ='NPP',
       x = 'Year', 
       y = unique(exp_object$NPP[scn == 'harvest_0', ][['unit']])) -> 
  NPP_plot

ggplot() + 
  geom_point(data = exp_object$LAI[scn == 'harvest_0', ], aes(datetime, value)) + 
  THEME + 
  labs(title ='LAI',
       x = 'Year', 
       y = unique(exp_object$LAI[scn == 'harvest_0', ][['unit']])) -> 
  LAI_plot

ggplot() + 
  geom_point(data = exp_object$NEP[scn == 'harvest_0', ], aes(year, value)) +
  THEME + 
  labs(title ='NEP',
       x = 'Year', 
       y = unique(exp_object$NEP[scn == 'harvest_0', ][['unit']])) -> 
  NEP_plot

ggplot() + 
  geom_point(data = exp_object$ABG[scn == 'harvest_0', ], aes(year, value)) + 
  THEME + 
  labs(title ='ABG',
       x = 'Year', 
       y = unique(exp_object$ABG[scn == 'harvest_0', ][['unit']])) -> 
  ABG_plot

plot_grid(ABG_plot, NPP_plot, NEP_plot, LAI_plot)

```

The base line scenario looks good to me it is intresting just how much smoother the output looks when we swtich to the climatological record instead of weather, do we want to only use the climatological values.

@@

So the differences in total above ground biomass between the different sencaiors is faily small and is really only visible when we zoom in on the disturbance and recovery. What is intresting is that the as we would expect the no harvest scenario has a larger biomass until September and then the AGB starts to even out, is pretty similar during winter but then with the next growing season the patches with the more extreme treatments have a larger biomass.

```{r}
exp_object$ABG$datetime <- date(exp_object$ABG$datetime)
data <- exp_object$ABG[  , value := sum(value), by = c("datetime", "scn", "unit", "year")]

data %>%  
  ggplot() + 
  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
  geom_point(aes(datetime, value, color = scn, group = interaction(scn, pft_name))) + 
  THEME + 
  labs(x = 'Year', y = unique(exp_object$ABG$unit), title = 'ABG') 
```

```{r}
data[year %in% 2018:2020 & scn != 'harvest_0', ] %>% 
  .[datetime >= date('2019-05-01')] %>%  
  ggplot() + 
  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
  geom_line(aes(datetime, value, color = scn, group = interaction(scn, pft_name)), size = 1) + 
  THEME + 
  labs(x = 'Year', y = unique(exp_object$ABG$unit), title = 'ABG') 
```

@@

```{r}
exp_object$NPP %>% 
  .[year %in% 2018:20205, ] %>% 
  ggplot() + 
  geom_vline(aes(xintercept = 2019), color ='black') + 
  geom_line(aes(year, value, color = scn, group = interaction(scn, pft_name))) + 
  THEME + 
  labs(x = 'Year', y = unique(exp_object$NPP$unit), title = 'Annual NPP')
```

@@

```{r}
h0_object  <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'test', 'baseline.rds'))
h45_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_45.rds'))
h65_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_65.rds'))
h85_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_85.rds'))

@@

h0$scn  <- 'harvest_0'
h45$scn <- 'harvest_45'
h65$scn <- 'harvest_65'
h85$scn <- 'harvest_85'

NPP <- rbind(h0, h45, h65, h85)
# Convert the NPP units from per plant to unit area
NPP$value <- NPP$MMEAN_NPP_CO * NPP$NPLANT
NPP$unit  <- unique(exp_object$NPP$unit)
NPP <- NPP[, c("datetime", "PFT", "scn", "value", 'unit')]
NPP$year <- lubridate::year(NPP$datetime)
NPP$datetime <- date(NPP$datetime)
NPP_monthly <- NPP 
```

```{r}
NPP_monthly <- NPP_monthly[, value := sum(value), by = c("datetime", "scn", "unit", "year")]

NPP_monthly %>% 
  ggplot(aes(datetime, value, color = scn)) + 
  # Note because of the fact that the results are recorded monthly the disturbance date is May 5th.
  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
  geom_line() + 
  THEME + 
  labs(title = 'Monthly NPP', y = unique(to_plot_NPP$unit), x = 'Time') 
```

```{r}
NPP_monthly %>% 
  .[year %in% 2018:2030] %>%  
  ggplot(aes(datetime, value, color = scn)) + 
  # Note because of the fact that the results are recorded monthly the disturbance date is May 5th.
  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
  geom_line() + 
  THEME + 
  labs(title = 'Monthly NPP', y = unique(to_plot_NPP$unit), x = 'Time') 
```

Would it make sense to look at the differrence between the baseline and treatment scenarios?

Observations

* Even after 12 years the stand has not recovered in term of NPP values 
* Also check out the difference in the winter NPP! I wasn't expecting that

@@

```{r}
exp_object$NEP %>% 
  dplyr::filter(year >= 2017 & year <= 2021) %>% 
  ggplot() + 
  geom_vline(xintercept = 2019, color = 'grey', size = 2) + 
  geom_line(aes(year, value, color = scn)) + 
  THEME + 
  labs(x = 'Year', y = unique(exp_object$NEP$unit), title = 'Annual NEP')
```

```{r}
h0_object  <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'test', 'baseline.rds'))
h45_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_45.rds'))
h65_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_65.rds'))
h85_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_85.rds'))

h0 <- as.data.table(h0_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]
h45 <- as.data.table(h45_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]
h65 <- as.data.table(h65_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]
h85 <- as.data.table(h85_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]

h0$scn  <- 'harvest_0'
h45$scn <- 'harvest_45'
h65$scn <- 'harvest_65'
h85$scn <- 'harvest_85'

NEP <- rbind(h0, h45, h65, h85)
NEP$unit  <- unique(exp_object$NEP$unit)
NEP <- NEP[, c("datetime", "scn", "value" = "MMEAN_NEP_PY")]
names(NEP) <- c("datetime", "scn", "value")
NEP$year <- lubridate::year(NEP$datetime)
NEP$datetime <- date(NEP$datetime)
```

```{r}
NEP %>% 
  dplyr::filter(year >= 2019 & year <= 2021) %>% 
  ggplot() + 
  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
  geom_line(aes(datetime, value, color = scn)) + 
  THEME + 
  labs(x = 'Year', y = unique(exp_object$NEP$unit), title = 'Monthly NEP')
```

@@

```{r}
exp_object$LAI$year <- year(exp_object$LAI$datetime)
exp_object$LAI %>%
  dplyr::filter(year >= 2019 & year <= 2021) %>% 
  ggplot() + 
  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
  geom_line(aes(datetime, value, color = scn)) + 
  THEME + 
  labs(x = 'Year', y = unique(exp_object$LAI$unit), title = 'Monthly LAI') + 
  facet_wrap('pft_name', ncol = 1)
```

@@

h0$scn  <- 'harvest_0'
h45$scn <- 'harvest_45'
h65$scn <- 'harvest_65'
h85$scn <- 'harvest_85'

@@

RH %>% 
ggplot(aes(datetime, MMEAN_RH_PY, color = scn)) + 
  geom_line() + 
  THEME + 
  labs(title = 'Monthly Heterotrophic Respiration', 
       y = units,
       x = 'Date Time')
```

```{r}
RH[ , year := year(datetime)] %>% 
  .[year %in% 2018:2020, ] %>% 
  ggplot(aes(datetime, MMEAN_RH_PY, color = scn)) + 
  geom_line() + 
  THEME + 
  labs(title = 'Monthly Heterotrophic Respiration', 
       y = units,
       x = 'Date Time')
```

## Mortality

Honestly, I don't really know what the morality output is so might not be easy to interpret but it looks like by 2020 the morality related to the disturbance events has ended.

```{r}
h0 <- as.data.table(h0_object$df_cohort[[1]])[  ,list(datetime, mmean_mort_rate_co_carbon)]
h45 <- as.data.table(h45_object$df_cohort[[1]])[  ,list(datetime, mmean_mort_rate_co_carbon)]
h65 <- as.data.table(h65_object$df_cohort[[1]])[  ,list(datetime, mmean_mort_rate_co_carbon)]
h85 <- as.data.table(h85_object$df_cohort[[1]])[  ,list(datetime, mmean_mort_rate_co_carbon)]

h0$scn  <- 'harvest_0'
h45$scn <- 'harvest_45'
h65$scn <- 'harvest_65'
h85$scn <- 'harvest_85'

mortality    <- rbind(h0, h45, h65, h85)
units <- '1/yr'

mortality[ ,year := year(datetime)] %>% 
  .[year %in% 2018:2021, ] %>% 
ggplot(aes(datetime, mmean_mort_rate_co_carbon, color = scn)) + 
  geom_line() + 
  THEME + 
  labs(title = 'Monthly Mean Mortlaity Rate per Cohort', 
       y = units,
       x = 'Date Time')
```

>>>>

---
title: "Experiment 1"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook: 
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    theme: lumen
---

@@

In experiment 1 we apply the different FoRTE girdling treatments to the ED forest stand starting from a near bare ground run starting in 1900.

@@

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# The defaults for the chunks
knitr::opts_chunk$set(echo = TRUE, fig.width = 12, fig.height = 12)

library(ggplot2)
library(magrittr)
library(data.table)
library(cowplot)
library(lubridate)
library(purrr)
library(fortedata)

BASE_DIR <- here::here()
THEME    <- theme_bw(base_size = 30)

ALPHA <- 0.05
BCK_GRND1 <- '#dcffcc'
BCK_GRND2 <- '#9fdfcd'
BCK_GRND3 <- '#baabda'
DISTURBANCE_COLORS <- c("0" = "#000000", "45" = "#009E73", "65" = "#0072B2", "85" = "#D55E00")

MONTHLY_BACKGROUND <- function(plot){
  plot + 
    geom_rect(aes(xmin=ymd('1900-01-01'), xmax = ymd('1979-12-01'),
                  ymin = -Inf, ymax = Inf), fill = BCK_GRND1, alpha = ALPHA) +
    geom_rect(aes(xmin=ymd('1980-01-01'), xmax = ymd('2015-12-01'),
                  ymin = -Inf, ymax = Inf), fill = BCK_GRND2, alpha = ALPHA) +
    geom_rect(aes(xmin=ymd('2016-01-01'), xmax = ymd('2029-01-01'),
                  ymin = -Inf, ymax = Inf), fill = BCK_GRND3, alpha = ALPHA) + 
    labs(caption = 'colored background: looped over 1979 - 1978, obs. 1979 - 2016, avg 2008 - 2018') 
}
ANNUAL_BACKGROUND <- function(plot){
  plot + 
    geom_rect(aes(xmin=1900, xmax = 1979,
                  ymin = -Inf, ymax = Inf), fill = BCK_GRND1, alpha = ALPHA) +
    geom_rect(aes(xmin=1979, xmax = 2015,
                  ymin = -Inf, ymax = Inf), fill = BCK_GRND2, alpha = ALPHA) +
    geom_rect(aes(xmin=2015, xmax = 2029,
                  ymin = -Inf, ymax = Inf), fill = BCK_GRND3, alpha = ALPHA) + 
    labs(caption = 'colored background: looped over 1979 - 1978, obs. 1979 - 2016, avg 2008 - 2018')
}
ADD_BACKGROUND <- function(plot){

  if(any(c('date', 'datetime') %in% names(plot$data))){
    p <- MONTHLY_BACKGROUND(plot)
  } else if ('year' %in% names(plot$data)){
    p <- ANNUAL_BACKGROUND(plot)
  }
  return(p)
}

ADD_DISTURBANCE_LINE_MONTH  <- geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2)
ADD_DISTURBANCE_LINE_ANNUAL <- geom_vline(xintercept = 2019, color = 'grey', size = 2)
ADD_COLORS <- scale_color_manual(values = DISTURBANCE_COLORS) 
```

@@

```{r}
# Note that this is the object created with the drake frame work, 
# there should be some better way to trigger this.
exp_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', "exp-1.rds"))
exp_object <- map(exp_object, function(x){ x[year < 2030, ] })

# Update the data for with the disturbance severity
add_disturbance_severity <- function(list){
  y <- map(list, function(x){
    x[data.table(scn = c("harvest 0", "harvest 45", "harvest 65", "harvest 85"), 
                 `disturbance severity` = c("0", "45", "65", "85")), on = 'scn'] 
  })
  return(y)
}

exp_object <- add_disturbance_severity(exp_object)
```

@@

# Meteorological Inputs

> 
* dlwrf — downward long wave radiation (W m-2)
* nbdsf — near infrared beam downward solar radiation (W m-2)
* nddsf — near IR diffuse downward solar radiation (W m-2)
* vbdsf — visible beam downward solar radiation (W m-2)
* vddsf — visible diffuse downward solar radiation (W m-2)
* prate — precipitation rate (kgH2O m-2 s-1)
* pres — atmospheric pressure (Pa)
* hgt — geopotential height (m)
* ugrd — zonal wind (m s-1)
* vgrd — meridional wind (m s-1)
* sh — specific humidity (kgH2O kgAir-1)
* tmp — air temperature (K)

```{r, fig.width = 12, fig.height = 6}
met_data <- as.data.table(read.csv(file.path(BASE_DIR, 'A.inputs', 'monthly_mean_met.csv')))
met_data$date <- ymd(paste(met_data$year, met_data$month, '01', sep = '/'))

long_met_data <- melt(met_data, measure.vars = c("nbdsf", "nddsf", "vbdsf", "vddsf",
                                                 "prate", "dlwrf", "pres", "hgt", "ugrd", 
                                                 "vgrd", "sh", "tmp"),
                      variable.name = 'variable', value.name = 'value')
long_met_data <- as.data.table(long_met_data[, c("date", "variable", "value", "year")])

var_info <- data.table(variable = c('dlwrf', 'nbdsf', 'nddsf', 'vbdsf',
                                    'vddsf', 'prate', 'pres', 'hgt', 
                                    'ugrd', 'vgrd', 'sh', 'tmp'), 
                       description = c('downward long wave radiation', 
                                       'near infrared beam downward solar radiation', 
                                       'near IR diffuse downward solar radiation', 
                                       'visible beam downward solar radiation',
                                       'visible diffuse downward solar radiation', 
                                       'precipitation rate', 
                                       'atmospheric pressure', 
                                       'geopotential height', 
                                       'zonal wind',
                                       'meridional wind', 
                                       'specific humidity',
                                       'air temperature'),  
                       units = c('W m-2', 'W m-2', 'W m-2', 'W m-2',
                                 'W m-2', 'kgH2O m-2 s-1', 'Pa', 'm',
                                 'm s-1', 'm s-', 'kgH2O kgAir-1', 'K'))
long_met_data <- long_met_data[var_info, on = 'variable']
long_met_data <- long_met_data[variable != 'hgt', ] # exclude the geo spatial height on cause it doesn't 
```

```{r}
ADD_BACKGROUND(plot = ggplot(data = long_met_data)) +
  geom_line(aes(date, value)) + 
  facet_wrap("variable", scales = 'free') + 
  labs(title = 'Monthly Time Series of ED Met Data Inputs', y = NULL) + 
  THEME
```

Because of the variability in the seasonal cycle is rather large and hides longer trem trends. Let's take a look at the annual min, max and mean.

```{r}
annual_data <- long_met_data[ , list('value' = mean(value), 'min' = min(value), 'max' = max(value)),
                              by = list(year, variable)]

ADD_BACKGROUND(plot = ggplot(data = annual_data)) +
  geom_ribbon(aes(year, ymin = min, ymax = max), fill = 'grey') + 
  geom_line(aes(year, min), color = 'grey') + 
  geom_line(aes(year, max), color = 'grey') + 
  geom_line(aes(year, value)) +
  facet_wrap('variable', scales = 'free') + 
  labs(y = NULL) 
```

Zooming in what does the actual observational data look like?

```{r}
ggplot(data = annual_data[year %in% 1979:2008, ]) +
  geom_ribbon(aes(year, ymin = min, ymax = max), fill = 'grey') + 
  geom_line(aes(year, min), color = 'grey') + 
  geom_line(aes(year, max), color = 'grey') + 
  geom_line(aes(year, value)) +
  facet_wrap('variable', scales = 'free') + 
  THEME + 
  labs(y = NULL)
```

@@

```{r, fig.width = 12, fig.height = 6}
# Process the data to plot, the annual values. 
exp_object$NPP[scn == 'harvest 0', ] %>%  
    .[  , value := sum(value), by = c('year', 'variable', 'unit', 'scn')] %>% 
    .[ , list(year, variable, unit, scn, value)] %>% 
  unique() -> 
  NPP_to_plot

exp_object$LAI[scn == 'harvest 0', ] %>%
  .[  , value := sum(value), by = c('datetime', 'variable', 'unit', 'scn')] %>% 
  .[ , list(datetime, variable, unit, scn, value)] %>%  
  unique() -> 
  LAI_to_plot

exp_object$NEP[scn == 'harvest 0', ] %>%  
  .[  , value := sum(value), by = c('year', 'variable', 'unit', 'scn')] %>% 
  .[ , list(year, variable, unit, scn, value)] %>%  
  unique() -> 
  NEP_to_plot

exp_object$ABG[scn == 'harvest 0', ] %>%  
  .[  , value := sum(value), by = c('year', 'variable', 'unit', 'scn')] %>% 
  .[ , list(year, variable, unit, scn, value)] %>%  
  unique() -> 
  ABG_to_plot

exp_object$GPP[scn == "harvest 0", ] %>%  
  .[  , value := sum(value), by = c('year', 'variable', 'unit', 'scn')] %>% 
  .[ , list(year, variable, unit, scn, value)] %>%  
  unique() -> 
  GPP_to_plot

ADD_BACKGROUND(plot = ggplot(data = NPP_to_plot)) + 
  geom_line(aes(year, value), size = 1) +
  THEME +
  labs(title ='NPP',
       x = 'Year', 
       y = unique(exp_object$NPP[scn == 'harvest 0', ][['unit']]), caption = NULL) -> 
  NPP_plot

ADD_BACKGROUND(plot = ggplot(data = LAI_to_plot)) + 
  geom_point(aes(datetime, value)) + 
  THEME + 
  labs(title ='LAI',
       x = 'Year', 
       y = unique(exp_object$LAI[scn == 'harvest 0', ][['unit']]), caption = NULL, size = 0.01) -> 
  LAI_plot

ADD_BACKGROUND(ggplot(NEP_to_plot)) + 
  geom_line(aes(year, value), size = 1) +
  THEME + 
  labs(title ='NEP',
       x = 'Year', 
       y = unique(exp_object$NEP[scn == 'harvest 0', ][['unit']]), caption = NULL) -> 
  NEP_plot

ADD_BACKGROUND(ggplot(ABG_to_plot)) + 
  geom_line(aes(year, value), size = 1) + 
  THEME + 
  labs(title ='ABG',
       x = 'Year', 
       y = unique(exp_object$ABG[scn == 'harvest 0', ][['unit']]), caption = NULL) -> 
  ABG_plot

ADD_BACKGROUND(ggplot(GPP_to_plot)) + 
  geom_point(aes(year, value), size = 1) + 
  THEME + 
  labs(title = 'GPP', 
       x = 'Year', 
       y = 'kg/m2/yr') -> 
  GPP_plot

plot_grid(ABG_plot, NPP_plot, NEP_plot, LAI_plot, GPP_plot)
```

One potential idea would be to use an ensemble of climate data to create a an envelope of runs.

@@

```{r, fig.width = 12, fig.height = 6}
exp_object$ABG$datetime <- date(exp_object$ABG$datetime)
data <- exp_object$ABG[  , value := sum(value), 
                         by = c("datetime", "scn", "unit", "year", "disturbance severity")] 
ggplot(data = data) + 
  ADD_DISTURBANCE_LINE_MONTH + 
  geom_line(aes(datetime, value, color = `disturbance severity`, 
                group = interaction(scn, pft_name)),
            size = 1.5) + 
  THEME + 
  ADD_COLORS + 
  labs(x = 'Year', y = unique(exp_object$ABG$unit), title = 'ABG') 
```

@@

```{r,  fig.width = 12, fig.height = 6}
ggplot(exp_object$NPP[year %in% 2000:2030, ]) +
  ADD_DISTURBANCE_LINE_ANNUAL +
  geom_line(aes(year, value, color = `disturbance severity`, 
                group = interaction(scn, pft_name)), size = 1.5) + 
  THEME + 
  coord_cartesian(xlim = c(2000, 2025)) + 
  labs(x = 'Year', y = unique(exp_object$NPP$unit), title = 'Annual NPP') + 
  ADD_COLORS
```

@@

```{r}
h0_object  <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_0_above.rds'))
h45_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_45_above.rds'))
h65_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_65_above.rds'))
h85_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_85_above.rds'))

@@

h0$severity  <- '0'
h45$severity <- '45'
h65$severity <- '65'
h85$severity <- '85'

NPP <- rbind(h0, h45, h65, h85)
# Convert the NPP units from per plant to unit area
NPP$value <- NPP$MMEAN_NPP_CO * NPP$NPLANT
NPP$unit  <- unique(exp_object$NPP$unit)
NPP <- NPP[, c("datetime", "PFT", "severity", "value", 'unit')]
NPP$year <- lubridate::year(NPP$datetime)
NPP$datetime <- date(NPP$datetime)
monthly_NPP <- NPP
```

```{r}
NPP_monthly <- monthly_NPP[, value := sum(value), by = c("datetime", "severity", "unit", "year")]
NPP_monthly$value <- NPP_monthly$value / 12

ggplot(NPP_monthly) + 
  ADD_DISTURBANCE_LINE_MONTH + 
  geom_line(aes(datetime, value, color = severity)) + 
  THEME + 
  ADD_COLORS +
  labs(title = 'NPP', y = 'kgC/m2/month', x = 'Time') 
```

```{r}
NPP_monthly %>% 
  .[year %in% 2018:2022] %>%  
  ggplot(aes(datetime, value, color = severity)) + 
  ADD_DISTURBANCE_LINE_MONTH +
  ADD_COLORS +
  geom_line(size = 1.5) + 
  THEME + 
  labs(title = 'Monthly NPP', y = 'kgC/m2/month', x = 'Time') 
```

@@

```{r}
exp_object$NEP %>% 
  dplyr::filter(year >= 2000 & year <= 2021) %>% 
  ggplot() + 
  geom_vline(xintercept = 2019, color = 'grey', size = 1.5) + 
  geom_line(aes(year, value, color = `disturbance severity`), size = 1.5) + 
  THEME + 
  ADD_COLORS +
  labs(x = 'Year', y = unique(exp_object$NEP$unit), title = 'Annual NEP')
```

```{r}
h0 <- as.data.table(h0_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]
h45 <- as.data.table(h45_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]
h65 <- as.data.table(h65_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]
h85 <- as.data.table(h85_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]

h0$severity  <- '0'
h45$severity <- '45'
h65$severity <- '65'
h85$severity <- '85'

NEP <- rbind(h0, h45, h65, h85)
NEP$unit  <- unique(exp_object$NEP$unit)
NEP <- NEP[, c("datetime", "severity", "value" = "MMEAN_NEP_PY")]
names(NEP) <- c("datetime", "severity", "value")
NEP$year <- lubridate::year(NEP$datetime)
NEP$datetime <- date(NEP$datetime)
```

```{r}
NEP %>% 
  dplyr::filter(year >= 2019 & year <= 2021) %>% 
  dplyr::mutate(value = value / 12, 
                units = 'kg/m2/month') %>% 
  ggplot() + 
  ADD_DISTURBANCE_LINE_MONTH + 
  geom_line(aes(datetime, value, color = severity), size = 1) + 
  THEME + 
  ADD_COLORS +
  labs(x = 'Year', y ='kg/m2/month', title = 'Monthly NEP')
```

@@

```{r}
exp_object$LAI$year <- year(exp_object$LAI$datetime)
exp_object$LAI %>% 
  dplyr::filter(year >= 2018 & year <= 2025) %>% 
  ggplot() + 
  ADD_DISTURBANCE_LINE_MONTH + 
  geom_line(aes(datetime, value, color = `disturbance severity`)) + 
  ADD_COLORS +
  THEME + 
  labs(x = 'Year', y = unique(exp_object$LAI$unit), title = 'Monthly LAI') + 
  facet_wrap('pft_name', ncol = 1)
```

@@

h0$severity  <- '0'
h45$severity <- '45'
h65$severity <- '65'
h85$severity <- '85'

@@

RH %>% 
ggplot(aes(datetime, MMEAN_RH_PY, color = severity)) + 
  geom_line() + 
  THEME + 
  labs(title = 'Monthly Heterotrophic Respiration', 
       y = units,
       x = 'Date Time') + 
  ADD_COLORS
```

```{r}
RH[datetime >= ymd('2018-06-01') & datetime <= ymd('2021-01-01'), ] %>% 
ggplot(aes(datetime, MMEAN_RH_PY, color = severity)) + 
  geom_line(size = 1.5) + 
  THEME + 
  labs(title = 'Monthly Heterotrophic Respiration', 
       y = units,
       x = 'Date Time') + 
  ADD_COLORS
```

## GPP

```{r}
GPP <- exp_object$GPP
GPP$severity <- GPP$`disturbance severity`

GPP %>% 
    .[  , value := mean(value), by = c('year', 'variable', 'unit', 'scn', 
                                      'PFT', 'description', 'severity', 
                                      'pft_name')] %>% 
    .[ , list(year, variable, scn, value, unit, pft_name,
              PFT, description, severity)] -> 
  GPP_to_plot

GPP_to_plot %>%    
  ggplot(aes(year, value, color = severity)) + 
  geom_point(size = 3) + 
  ADD_COLORS + 
  THEME + 
  labs(title = 'Annual GPP per Treatment', 
       y = 'kgC/m2/yr')
```

```{r}
GPP <- exp_object$GPP
GPP$severity <- GPP[['disturbance severity']]
GPP$datetime <- date(GPP$datetime)
GPP$PFT <- as.character(GPP$PFT)

# Add the cohort labels to the data
split(GPP, interaction(GPP$datetime, GPP$severity, GPP$PFT)) %>% 
  lapply(function(x){

    x$cohort <- LETTERS[1:nrow(x)]
    return(x)

  }) %>%  
  rlist::list.rbind() -> 
  GPP_cohort

to_plot <- GPP_cohort[year %in% 2018:2020, ] 
to_plot$value <- to_plot$value / 12
to_plot %>%  
  ggplot(aes(datetime, value, group = interaction(cohort, severity), color = severity)) + 
  geom_line(size = 2) + 
  facet_grid(. ~ pft_name) + 
  THEME + 
  ADD_COLORS + 
  labs(title = 'Cohort GPP', 
       y = 'kgC/m2/month', 
       x = 'Date')
```

## Carbon Cycling Resistance

Carbon Cycling Resistance

$$r = \ln \left( {\frac{{F_{{\text{dist}}} }}{{F_{{\text{cont}}} }}} \right)$$ Equation 1 from Gough 2020.

```{r}

exp_object$NPP[year %in% 2018:2030, ] %>%  
  .[, list(pft_name, year, variable, unit, value, severity = `disturbance severity`)] %>% 
  # Make the table wide by disturbance severity treatment. 
  dcast(data = ., pft_name + year + variable + unit ~ severity) %>%
  # Now gather all of the distubrance treatments into a single column so that the 
  # control values are recorded in their own column. 
  melt(data = ., id.vars = c("pft_name", "year", "variable", "unit", "0"), 
       measure.vars = c('45', '65', '85'), variable.name = 'severity', 
       value.name = 'treatment_value') %>%  
  # Rename the columns 
  .[ , list(pft_name, year, variable, unit, control = `0`, severity, dist = treatment_value)] %>% 
  # Now calculate r 
  .[ , r := log(dist/control)] -> 
  NPP_r

NPP_r %>%   
ggplot(aes(year, r, color = severity)) + 
  geom_line(size = 1.5) + 
  ADD_COLORS + 
  THEME + 
  labs(title = 'NPP Resistance Values', 
       y = expression('r'[npp]), 
       x = 'Year')
```
